---
title: "用tidyr清洁数据"
author: "张平"
date: "2019年4月26日"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
```
```{r}
table1 %>% 
  count(year, wt = cases)

ggplot(table1, aes(year, cases)) + 
  geom_line(aes(group = country), colour = "blue") + 
  geom_point(aes(colour = country))
```
## gather(): 宽表变成长表
```{r}
table4a
table4a %>% 
  gather(`1999`, `2000`, key = "year", value = "cases")
table4b
table4b %>% 
  gather(`1999`, `2000`, key = "year", value = "population")
tidy4a <- table4a %>% 
  gather(`1999`, `2000`, key = "year", value = "cases")
tidy4b <- table4b %>% 
  gather(`1999`, `2000`, key = "year", value = "population")
left_join(tidy4a, tidy4b)
```

## spread: 长表变为宽表
```{r}
table2
spread(data = table2, key = type, value = count)
```

### spread必须要有独特的标识符
```{r}
people <- tribble(
  ~name,             ~key,    ~value,
  #-----------------|--------|------
  "Phillip Woods",   "age",       45,
  "Phillip Woods",   "height",   186,
  "Phillip Woods",   "age",       50,
  "Jessica Cordero", "age",       37,
  "Jessica Cordero", "height",   156
)

people %>% mutate(id = row_number()) %>% spread(key = key, value = value)
```

## separate和unite
By default, separate() will split values wherever it sees a non-alphanumeric character (i.e. a character that isn’t a number or letter).
If you wish to use a specific character to separate a column, you can pass the character to the sep argument of separate().
You can also pass a vector of integers to sep. separate() will interpret the integers as positions to split at.
```{r}
table3 
table3 %>% separate(rate, into = c("cases", "population"))
table3 %>% separate(rate, into = c("cases", "population"), convert = TRUE)
table3 %>% 
  separate(year, into = c("century", "year"), sep = 2)
```
unite() is the inverse of separate(): it combines multiple columns into a single column.
The default will place an underscore (_) between the values from different columns. Here we don’t want any separator so we use "".
```{r}
table5 %>% 
  unite(new, century, year)
table5 %>% 
  unite(new, century, year, sep = "")
```

## 缺失值
Another important tool for making missing values explicit in tidy data is complete().
```{r}
stocks <- tibble(
  year   = c(2015, 2015, 2015, 2015, 2016, 2016, 2016),
  qtr    = c(   1,    2,    3,    4,    2,    3,    4),
  return = c(1.88, 0.59, 0.35,   NA, 0.92, 0.17, 2.66)
)
stocks %>% 
  spread(year, return)

stocks %>% 
  spread(year, return) %>% 
  gather(year, return, `2015`:`2016`, na.rm = TRUE)

stocks %>% 
  complete(year, qtr)
stocks
```

fill in these missing values with fill(). It takes a set of columns where
you want missing values to be replaced by the most recent non-missing value
(sometimes called last observation.
```{r}
treatment <- tribble(
  ~ person, ~ treatment, ~response,
  "Derrick Whitmore", 1, 7,
  NA, 2, 10,
  NA, 3, 9,
  "Katherine Burke", 1, 4
)
treatment %>%
  fill(person)
```

